Hydrodynamics and rheology of weakly adhesive vesicle doublets

### 2019-02-04 vesicle_code/docs/adhesion/adhesion.tex
all the changes from yny so far

%
The electrostatic nature of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicates the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces,PerutkovaFrank-Bertoncelij2013_CSB} 
or a solid (such as a glass substrate
or a nano particle).  Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase,SteinkuhlerAgudo-Canalejo2016_BJ}
than the transient adhesion process \cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
%
Adhesive interactions between lipid membranes are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is crucial to initiate membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.
In the absence of an external electric field and ions in the solvents, 
it is reported that 
%such complicated electrostatic component in 
the adhesive interactions between two membranes can be well approximated by the 
L-J-type potential \cite{FlormannAouane2017_SciReports}, which consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. 
%%
%%move this somewhere else Jan30YNY
%Such adhesive
%potentials are essential in many biomedical, biological and biophysical
%processes.  For example, vesicle adhesion is essential for membrane
%fusion/fission in endocytosis and exocytosis, the transport of small
%vesicles through membrane surfaces.
%%move this somewhere else Jan30YNY

%Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ} used the micropipette to measure the adhesive force between two vesicles bound by strong adhesion.
They were able to  measure the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their Van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong Van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\it et al.} used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

@@

Next we conduct a numerical experiment where a vesicle doublet is placed at the center of a fluid trap, which in microfluidic laboratory 
can be actively controlled so that fluid trap center is effectively a stagnation point in an extensional flow characterized by a extensional flow rate.
At low flow rate, we find the doublet to rotate ninety degrees to align with the flow such that 
the long axis of the doublet is along the divergent axis and the convergent stream is pushing the two vesicles
together. As the flow rate increases the vesicle doublet rotates less and we observe tank-treading like motion of both vesicles in the doublet.
When the flow rate reaches above the critical value, the diverging flow pulls the vesicles apart before the doublet has time to rotate to remain intact.
These results indicate that it is possible to use the fluid trap to separate a vesicle doublet under adhesion, and thus provide a means to
probe the adhesion strength between membranes. {\color{red} For adhesion energy blahblahblah, an extensional flow rate sososo is required to separate the doublet from our
simulations. IS THIS PLAUSIBLE?}
We hope that our simulations might inspire microfluidic experiments.

>>>>

%
The electrostatic nature of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicates the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces,PerutkovaFrank-Bertoncelij2013_CSB} 
or a solid (such as a glass substrate
or a nano particle).  Adhesion between a vesicle and a substrate has been extensively studied with 
more focuses on the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, GruhnFrankeDimova2007_Langmuir,das2008adhesion, zhang2009phase,SteinkuhlerAgudo-Canalejo2016_BJ}
than the transient adhesion process \cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
%
Adhesive interactions between lipid membranes are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is crucial to initiate membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.
In the absence of an external electric field and ions in the solvents, 
it is reported that 
%such complicated electrostatic component in 
the adhesive interactions between two membranes can be well approximated by the 
L-J-type potential \cite{FlormannAouane2017_SciReports}, which consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. 
%%
%%move this somewhere else Jan30YNY
%Such adhesive
%potentials are essential in many biomedical, biological and biophysical
%processes.  For example, vesicle adhesion is essential for membrane
%fusion/fission in endocytosis and exocytosis, the transport of small
%vesicles through membrane surfaces.
%%move this somewhere else Jan30YNY

%Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ} used the micropipette to measure the adhesive force between two vesicles bound by strong adhesion.
They were able to  measure the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the van der Waals' attraction. For the adhesive interaction between two lipid bilayer membranes in buffer solutions, a typical range for the
energy density is between $1$ to $10$ $\mu J/m^2$, 
%
%they reported
%an energy density of $\sim 10$  $\mu J/m^2$ and a separation distance of $\sim 3$  $nm$.
%
% membrane-membrane
%separation distance of $\sim 3$ $nm$, they estimated that the Hamaker
%constant is approximately $5.8\times 10^{-14}$ $ergs$, which corresponds
%to 
%
%an adhesion energy density around $1$ $\mu J/m^2$,
similar to the adhesion energy density of $3$  $\mu J/m^2$ between two RBCs \cite{FlormannAouane2017_SciReports}.
For adhesion interaction between a vesicle and a substrate  \cite{GruhnFrankeDimova2007_Langmuir}, 
the boundary between weak and strong adhesion is around $1$ $\mu J/m^2$:
the vesicle-substrate interaction is ``strong" when the adhesion energy density is larger than $1$ $\mu J/m^2$ and ``weak" when the energy
density is less than $1$ $\mu J/m^2$.
%
%weak adhesion regime 
%
%Gruhn {\it et al.} made a distinction between strong and weak adhesion regimes for a vesicle adhering on a substrate.
%Strong adhesion regime when the adhesion energy density is greater than $1$ $\mu J/m^2$,
%while weak adhesion regime when adhesion energy density is less than $1$ $\mu J/m^2$.
%
In this work we use the ratio of the adhesion energy to the bending energy for classification of adhesion strength:
The ratio of total adhesion
energy to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ $J$) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.

In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (adhesion energy/bending energy $\ll 1$) and strong adhesion (adhesion energy/bending energy $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes in the strong adhesion regime
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir,SteinkuhlerAgudo-Canalejo2016_BJ,FlormannAouane2017_SciReports} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\it et al.} used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

@@

Next we conduct a numerical experiment where a vesicle doublet is placed at the center of a fluid trap, which 
can be actively controlled in microfluidic channel so that fluid trap center is effectively the stagnation point of an extensional flow characterized by a extensional flow rate.
At low flow rate, we find the doublet to rotate ninety degrees to align with the flow such that 
the long axis of the doublet is along the divergent axis and the convergent stream is pushing the two vesicles
together. As the flow rate increases the vesicle doublet rotates less and we observe tank-treading like motion of both vesicles in the doublet.
When the flow rate reaches above the critical value, the diverging flow pulls the vesicles apart before the doublet has time to rotate to remain intact.
These results indicate that it is possible to use the fluid trap to separate a vesicle doublet under adhesion, and thus provide a means to
probe the adhesion strength between membranes. 
For a pair of $\mu m$-sized vesicles with a bending rigidity of $\sim 10^{-19}$ $J$, an extensional flow rate of $0.5$ $s^{-1}$ is expected to separate a vesicle
doublet with reduced area of $0.8$ and a Hamakar constant ${\cal H}=0.7$, which corresponds to $\sim 1$ $\mu J/m^2$. These conditions are quite realizable in micrifulidic
experiments, and we hope that our simulations might inspire microfluidic experiments.
%{\color{red} For adhesion energy blahblahblah, an extensional flow rate sososo is required to separate the doublet from our
%simulations. IS THIS PLAUSIBLE?}
